\Title

### 2021-05-28 latex/paper/main.tex
Partway through the proof of Thm 10, still working out the best way to do this.

%/========== Proof ==========/%
\section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:proof}
\textbf{TODO: prove Thm 10}
The sets \(\mathcal{O}_1\) and \(\mathcal{O}_2(\bar{\rho}\) are sublevel sets of
the nominal energy \(E(q_u,p_u)\) defined by \eqref{eqn:acrobot-energy}.
To prove Theorem \ref{thm:acrobot-energy-stabilization}, we will find conditions
under which the acrobot gains energy on an open ``energy annulus".
That is, given two energy levels \(E_0 \geq 0\) and \(E_1 > E_0\), we define the
energy annulus to be the domain
\[
    \mathcal{D} := \{(q_u,p_u) \in \SxR \mid \\
    E_0 < E(q_u,p_u) < E_1\}
    .
\]
The only condition we place on \(E_0\) and \(E_1\) is that they both correspond
to oscillations or both to rotations; that is, if \(E_0 < E(\pi,0)\), then 
\(E_1 \leq E(\pi,0)\).

@@

We will describe the explicit forms of these transformations later. 
Thanks to the shape of a pendulum's orbits, we will show that (at \(I = 0\)) the
dynamics of \((r,\theta)\) satisfy \(\dot{r} = 0\) and \(\dot{\theta} > 0\).
By continuity of the constrained dynamics with respect to \(I\), we can find a
small enough \(I > 0\) so that \(\dot{\theta} > 0\) for the acrobot.
This property will enable us to perform a time-scaling transformation 
\(t = t(\theta)\) to eliminate the dynamics of \(\theta\), and study how 
the pseudo-radius \(r\) evolves as a function of the pseudo-angle \(\theta\),
rather than as a function of time.
If it increases ``on average" , then \(E\) must also be increasing ``on average"
and the constrained dynamics must be gaining energy on \(D\).

@@

\begin{lemma}
    Consider the simplified acrobot \eqref{eqn:acrobot-hamiltonian} constrained
    by the VNHC \eqref{eqn:acrobot-constraint}, whose constrained dynamics are
    parameterized by \((q_u,p_u) \in \SxR\).
    Let \(D \subset \SxR\) be open and 
    let \(\underline{r}, \overline{r} \in \R_{> 0}\) with 
    \(\underline{r} < \overline{r}\).
    Suppose there exists a diffeomorphism 
    \begin{align*}
        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone \\
        (q_u,p_u) &\mapsto (r,\theta)
        ,
    \end{align*}
    satisfying \(\dot{r} = 0\) and \(\dot{\theta} > 0\) at \(I = 0\).
    Let \(g(\theta,r,I) = \dot{r}/\dot{\theta}\) and assume that 
    \((\partial g/\partial I)(\theta,r,0) = 0\).
    Define \(b(\theta,r_0) = g(\theta,r_0,0)\) and assume there exists
    \(\epsilon > 0\) such that 
    \[
        \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
        ,
    \]
    for all \(r_0 \in \, ]\underline{r},\overline{r}[\).
    Then there exists \(I^\star > 0\) small enough that the constrained dynamics
    gain energy on \(D\) for all \(0 < I \leq I^\star\), and lose energy on \(D\)
    for all \(-I^\star \leq I < 0\).
\end{lemma}
%/========== Conclusion ==========/%
\section{Conclusion}\label{sec:conclusion}

>>>>

%/========== Proof ==========/%
\section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:proof}
The sets \(\mathcal{O}_1\) and \(\mathcal{O}_2(\bar{\rho})\) are sublevel sets of
the nominal pendulum's energy \(E(q_u,p_u)\) described in
\eqref{eqn:acrobot-energy}.
To prove Theorem \ref{thm:acrobot-energy-stabilization}, we will find conditions
under which the acrobot gains energy on an open ``energy annulus".
That is, given two energy levels \(E_0 \geq 0\) and \(E_1 > E_0\), we define the
energy annulus to be the domain
\begin{equation*}
    \mathcal{D} := \{(q_u,p_u) \in \SxR \mid \\
    E_0 < E(q_u,p_u) < E_1\}
    .
\end{equation*}
The only condition we place on \(E_0\) and \(E_1\) is that they both correspond
to oscillations or both to rotations; that is, if \(E_0 < E(\pi,0)\), then 
\(E_1 \leq E(\pi,0)\).

@@

We will describe the explicit forms of these transformations later. 
Importantly, we will show that the dynamics of \((r,\theta)\) satisfy 
\(\dot{r} = 0\) and \(\dot{\theta} > 0\) for the nominal pendulum (i.e., when
\(I = 0\)).
By continuity of the constrained dynamics with respect to \(I\), we can find a
small enough \(I > 0\) so that \(\dot{\theta} > 0\) for the acrobot.
This property will enable us to perform a time-scaling transformation 
\(t = t(\theta)\), and study how the pseudo-radius \(r\) evolves as a function
of the pseudo-angle \(\theta\), rather than as a function of time.
If \(r(\theta)\) increases on average, then \(E\) must also be increasing
on average and the constrained dynamics must be gaining energy on \(D\).

@@

\begin{lemma}
    Consider the simplified acrobot \eqref{eqn:acrobot-hamiltonian} constrained
    by the VNHC \eqref{eqn:acrobot-constraint}, whose constrained dynamics are
    parameterized by \((q_u,p_u) \in \SxR\).
    Let \(D \subset \SxR\) be open and 
    let \(\underline{r}, \overline{r} \in \R_{> 0}\) with 
    \(\underline{r} < \overline{r}\).
    Suppose there exists a diffeomorphism 
    \begin{align*}
        T : D &\to T(D) = ]\underline{r},\overline{r}[ \, \times \Sone \\
        (q_u,p_u) &\mapsto (r,\theta)
        ,
    \end{align*}
    satisfying \(\dot{r} = 0\) and \(\dot{\theta} > 0\) on \(T(D)\) when 
    \(I = 0\).
    Let \(g(\theta,r,I) = \dot{r}/\dot{\theta}\) and assume that 
    \((\partial g/\partial I)(\theta,r,0) = 0\).
    Define \(b(\theta,r_0) = g(\theta,r_0,0)\) and assume there exists
    \(\epsilon > 0\) such that 
    \[
        \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
        ,
    \]
    for all \(r_0 \in \, ]\underline{r},\overline{r}[\).
    Then there exists \(I^\star > 0\) small enough that the constrained dynamics
    gain energy on \(D\) for all \(0 < I \leq I^\star\), and lose energy on \(D\)
    for all \(-I^\star \leq I < 0\).
\end{lemma}
\begin{proof}
    Since \(\dot{\theta} > 0\) when \(I = 0\) and the constrained dynamics 
    in \((r,\theta)\) coordinates are continuous in \(I\), there exists 
    \(I_1 > 0\) small enough that \(\dot{\theta} > 0\) for all
    \(I \in \, ]0,I_1]\).
    Hence, we can perform a time reparameterization \(t = t(\theta)\) to get a
    time-scaled pseudo-radius \(\hat{r}(\theta) := r(t(\theta))\).
    This reduces the system \((\dot{r},\dot{\theta})\) into the scalar
    time-varying ODE 
    \begin{equation}\label{eqn:rhat-ode}
        \begin{cases}
            \hat{r}^\prime= \frac{\dot{r}}{\dot{\theta}} =
            g(\theta,\hat{r},I)
            , \\
            \hat{r}(0) = r_0
            ,
        \end{cases}
    \end{equation}
    where the prime denotes differentiation by \(\theta\) and 
    \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.

    Let \(\hat{r}(\theta,r_0,I)\) be the solution to \eqref{eqn:rhat-ode}.
    Taking the Taylor expansion of this solution around \(I = 0\) we get
    \(\hat{r}(\theta,r_0,I) = \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0,0) 
    + R(\theta,r_0,I)\) where

    \textbf{TODO: finish the proof}
\end{proof}
\textbf{TODO: Prove origin is unstable, then apply lemma to O1 and O2 and show
that orbits go through Epi}
%/========== Conclusion ==========/%
\section{Conclusion}\label{sec:conclusion}
